Computational methods in landscape ecology

### 2024-10-20 Sections/04-connectivity.qmd
Move around some sentences

<!-- R1: I did not find the subsection "Structural connectivity" very consistent with the other two ("Resistance surface" and "Graph theory"). In addition to being very short, I wonder why this section was included without another on functional connectivity, and why it was distinguished from "Graph theory" (in terms of manuscript structure), since structural (as long as functional) connectivity can be modelled with graph theory. -->

<!-- R1: Here connectivity seems to be described through the prism of animals, you should modify it slightly to include plants, which do not move, but disperse through different vectors, e.g. animals, gravity, wind, or water -->
<!-- R1: facilitate or impeded movement and flow -->
Landscape connectivity describes how landscape features facilitate or impede movement, flow, and dispersal of organisms (e.g., active movement of animals, or dispersal of sessile plants by wind or water). 
It underpins ecosystem functionality, maintains biodiversity and populations, and plays an important role in many conservation actions [@CorreaAyram2016HabitatConnectivity].
Connectivity can be divided into structural and functional connectivity [@Taylor2006LandscapeConnectivity].
Structural connectivity describes the physical arrangement of landscape elements, focusing on spatial aspects such as continuity and adjacency.
It is solely a landscape feature and independent of species characteristics [@Tischendorf2000UsageMeasurement; @CorreaAyram2016HabitatConnectivity].
Landscape metrics are commonly used to describe structural connectivity in terms of, e.g., total habitat amount, patch size, or patch isolation.
Recently, there has been an increasing focus on within-patch connectivity using metrics such as the effective mesh size [@Spanowicz2019MeasuringLandscape].
However, landscape metrics are frequently criticized for an ambiguous link to functional connectivity [@Dramstad2009SpatialMetrics; @Kupfer2012LandscapeEcology].
Functional connectivity, in contrast, integrates landscape structure with the perceptual, behavioral, and dispersal characteristics of species, making it both species- and landscape-specific [@Adriaensen2003ApplicationLeastcost; @Taylor2006LandscapeConnectivity].

>>>>

<!-- R1: I did not find the subsection "Structural connectivity" very consistent with the other two ("Resistance surface" and "Graph theory"). In addition to being very short, I wonder why this section was included without another on functional connectivity, and why it was distinguished from "Graph theory" (in terms of manuscript structure), since structural (as long as functional) connectivity can be modelled with graph theory. -->
<!-- R1: Here connectivity seems to be described through the prism of animals, you should modify it slightly to include plants, which do not move, but disperse through different vectors, e.g. animals, gravity, wind, or water -->

<!-- R1: facilitate or impeded movement and flow -->
Landscape connectivity describes how landscape features facilitate or impede movement, flow, and dispersal of organisms (e.g., active movement of animals, or dispersal of sessile plants by wind or water). 
It underpins ecosystem functionality, maintains biodiversity and populations, and plays an important role in many conservation actions [@CorreaAyram2016HabitatConnectivity].
Connectivity can be divided into structural and functional connectivity [@Taylor2006LandscapeConnectivity].

Structural connectivity describes the physical arrangement of landscape elements, focusing on spatial aspects such as continuity and adjacency.
It is solely a landscape feature and independent of species characteristics [@Tischendorf2000UsageMeasurement; @CorreaAyram2016HabitatConnectivity].
Landscape metrics are commonly used to describe structural connectivity in terms of, e.g., total habitat amount, patch size, or patch isolation.
Recently, there has been an increasing focus on within-patch connectivity using metrics such as the effective mesh size [@Spanowicz2019MeasuringLandscape].
However, landscape metrics are frequently criticized for an ambiguous link to functional connectivity [@Dramstad2009SpatialMetrics; @Kupfer2012LandscapeEcology].
Functional connectivity, in contrast, integrates landscape structure with the perceptual, behavioral, and dispersal characteristics of species, making it both species- and landscape-specific [@Adriaensen2003ApplicationLeastcost; @Taylor2006LandscapeConnectivity].

### 2024-10-20 Sections/06-simulations.qmd
Move around some sentences

Individual-based models (or agent-based models) simulate discrete entities that are described by attributes and behavior.
Patterns emerge from bottom-up interactions of individuals with each other and their environment [@DeAngelis2014IndividualbasedModels].
In landscape ecology individual-based models are increasingly used to model social-ecological systems [@Carauta2018CanPreferential; @Dislich2018LanduseChange; @Gonzalez-Redin2019ExploringSustainable], but also disturbances [@Seidl2017ClimateChange; @Spies2017UsingAgentbased; @Kamimura2019AgentbasedModelling], or connectivity [@Day2020IndividualbasedModeling].
They can also incorporate a range of socio-economic, political, and governance information to influence how individuals, such as farmers, interact within a landscape [@Murray-Rust2014CombiningAgent].
In order to facilitate future model developments recent progress includes adaptable modeling frameworks [@Schumaker2018HexSimModeling] or reusable building blocks [@Berger2024ReusableBuilding].
Furthermore, hybrid approaches, coupling or integrating different model types with individual-based models, could benefit from rich development histories or facilitate corresponding strengths [@OSullivan2016StrategicDirections], e.g., linking an individual-based and ecosystem model [@Polhill2011AgentbasedModelling].

@@

The digital twins concept include three connected elements, namely *i)* a physical object or system, *ii)* its digital representation, *iii)* and a data exchange between the former two [@Jones2020CharacterisingDigital; @deKoning2023DigitalTwins].
Digital twins are particular useful for real-time workflows that allow now- and forecasting of complex dynamics in a landscape.
The need for timely and evidence-based decision making in combination with increasing data availability make digital twins a powerful tool for landscape ecology.
They aim to provide real-time state of nature measurements, early detection of conservation trends, relationships of ecosystem trends and environmental conditions, impact assessments of interventions, or to identify uncertainties and information gaps [@deKoning2023DigitalTwins].
Thus, there is increasing interest and use of digital twins in both the industry and academia [@Jones2020CharacterisingDigital].
Digital twins are increasingly applied in agricultural landscapes research, e.g., for livestock farming, controlled environment farming, or fertilization management [@Cesco2023SmartAgriculture; @Purcell2023DigitalTwins]. 
Further applications include planning of rural ecological landscapes [@Tan2024DigitalTwin] or exploration of relationships between urban expansion and vegetation coverage [@Zhao2022ApplyingDigital].
The use of general data processing, statistical and mechanistic models, or artificial intelligence allows digital twins to constantly update the digital representation and analyze, visualize, or predict the physical counterpart.
<!-- R1: You could also mention the increasing availability and accessibility of sensors deployed by drones along with data processing tools (e.g. drone photogrammetry, LiDAR) which enable high-frequency, high-resolution monitoring of 2D and 3D landscape models -->
This is further facilitated by the increasing rate of novel sources of data generation in ecology, e.g., due to sensors deployed on airplane, satellites, or unmanned aerial vehicle [@Farley2018SituatingEcology].

>>>>

Individual-based models (or agent-based models) simulate discrete entities that are described by attributes and behavior.
Patterns emerge from bottom-up interactions of individuals with each other and their environment [@DeAngelis2014IndividualbasedModels].
In landscape ecology individual-based models are increasingly used to model social-ecological systems [@Carauta2018CanPreferential; @Dislich2018LanduseChange; @Gonzalez-Redin2019ExploringSustainable], but also disturbances [@Seidl2017ClimateChange; @Spies2017UsingAgentbased; @Kamimura2019AgentbasedModelling], or connectivity [@Day2020IndividualbasedModeling].
They can also incorporate a range of socio-economic, political, and governance information to influence how individuals, such as farmers, interact within a landscape [@Murray-Rust2014CombiningAgent].

In order to facilitate future model developments recent progress includes adaptable modeling frameworks [@Schumaker2018HexSimModeling] or reusable building blocks [@Berger2024ReusableBuilding].
Furthermore, hybrid approaches, coupling or integrating different model types with individual-based models, could benefit from rich development histories or facilitate corresponding strengths [@OSullivan2016StrategicDirections], e.g., linking an individual-based and ecosystem model [@Polhill2011AgentbasedModelling].

@@

The digital twins concept include three connected elements, namely *i)* a physical object or system, *ii)* its digital representation, *iii)* and a data exchange between the former two [@Jones2020CharacterisingDigital; @deKoning2023DigitalTwins].
The need for timely and evidence-based decision making in combination with increasing data availability make digital twins a powerful tool for landscape ecology.
The use of general data processing, statistical and mechanistic models, or artificial intelligence allows digital twins to constantly update the digital representation and analyze, visualize, or predict the physical counterpart.
<!-- R1: You could also mention the increasing availability and accessibility of sensors deployed by drones along with data processing tools (e.g. drone photogrammetry, LiDAR) which enable high-frequency, high-resolution monitoring of 2D and 3D landscape models -->
This is further facilitated by the increasing rate of novel sources of data generation in ecology, e.g., due to sensors deployed on airplane, satellites, or unmanned aerial vehicle [@Farley2018SituatingEcology].
Digital twins are particular useful for real-time workflows that allow now- and forecasting of complex dynamics in a landscape.

Digital twins aim to provide real-time state of nature measurements, early detection of conservation trends, relationships of ecosystem trends and environmental conditions, impact assessments of interventions, or to identify uncertainties and information gaps [@deKoning2023DigitalTwins].
Thus, there is increasing interest and use of digital twins in both the industry and academia [@Jones2020CharacterisingDigital].
They  are increasingly applied in agricultural landscapes research, e.g., for livestock farming, controlled environment farming, or fertilization management [@Cesco2023SmartAgriculture; @Purcell2023DigitalTwins]. 
Further applications include planning of rural ecological landscapes [@Tan2024DigitalTwin] or exploration of relationships between urban expansion and vegetation coverage [@Zhao2022ApplyingDigital].
